### 2019-08-31 Writeup/Paper/sections/background.tex
Changes; - Changes made to the precis paper structure

\section{Background}
In this section we dicuss

@@

\subsection{Similarity Metrics}
\label{sec:similarity_metric}
A similarity metric is an algorithm designed to determine if two words are phonetically a match. For example, the words "THEIR" and "THERE" is a match, whereas the words "DARK" and "PRINCIPLE" are not phonetically matching. This section provides a base explanation of important phonetic algorithms relevant in this project.

\subsubsection*{Soundex}
\label{sec:soundex}
One of the earliest examples of a phonetic algorithm is known as Soundex. It was initially designed for phonetically indexing names alongside detection of transposed letters in spelling mistakes.  Soundex is one of the most famous example of a phonetic. This is due in part to its implementation into major database clients like MySQL\cite{mysql_soundex}, Oracle\cite{moved_2005} and PostgreSQL\cite{postgresql}.

\subsubsection*{NYSIIS}
\label{sec:nysiis}
The New York State Identification and Intelligence System (NYSIIS) phonetic code was created for the phonetic matching of American names. The motivation for its inception was mostly due to the presence of Hispanic names in the American based databases (this was an aspect Soundex was known to have low accuracy with).

\subsubsection*{Metaphone}
\label{sec:metaphone}
Metaphone was invented by Lawrence Philips in 1990\cite{philips1990hanging} in response to the deficiencies in Soundex. It improves on Soundex by including information around inconsistency and variation in English spelling in an attempt to create a more accurate phonetic representation.

\subsubsection*{Levenshtein Distance}
\label{sec:leven}
Levenshtein distance is a string metric designed to measure the `distance' between two strings. It is merely the number of single-character edits (insertions, deletions or substitutions) required to reach the other string.

An example distance between \verb|trace| and \verb|place| would be the substitutions of the first to letters, from \verb|tr| to \verb|pl|, meaning the two strings have a Levenshtein distance of 2.

\subsubsection*{Phonetic Vectors}
\label{sec:phonetic_vectors}
Phonetic Vectors is a unique addition to the chosen set. Created by Allison Parrish in 2017\cite{parrish2017poetic}, Phonetic vectors is as the name suggests the vectorisation of the phonetics of a word. Phonetic features are used in this work as a way to compare the similarity of phonemes. Phonemes are the phonetic elements that construct a word. For example, the word "RING" translated into the phonemes \verb|/R IH NG/|.

\clearpage

\subsection{Pretty Easy Privacy}
\label{sec:pep}
A very recent implementation of a word list can be found in Pretty Easy Privacy (\pep) implementation of TrustWords\cite{trustwords}. \pep is a data encryption system that utilises PGP encryption to provide E2EE on all common channels of communication such as email or SMS. The embedded design principles state that above all the systems should be easy to install, use and understand.

\pep deals with the threat of MiTM attacks by having users compare the respective key fingerprints encoded as a set of words. Figure \ref{fig:trustwords} shows the \pep Android implementation of Trustwords. The users then authenticate the words on an OOB (Out of Band) channel such as a phone call or in-person communication. If both users decide the words match, they accept or decline respectively.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.4\textwidth]{trustwords/trustword_handshake.jpg}
    \caption{Trustword fingerprint verification}
    \label{fig:trustwords}
\end{figure}

The unique aspect of TrustWords is its mapping of a single word to 16-bits. If compared to alternative literature, this is the highest number of bits-per-word seen. Full mappings (no duplication of words) would, therefore, require $2^{16}$ words in the dictionary. This number of words is arguably higher than most users' vocabulary. This deviation from the norm has not been currently backed up by research.

>>>>

\section{Background}

@@

\subsection{Similarity Metrics}
\label{sec:similarity_metric}
A similarity metric is an algorithm designed to determine if two words are phonetically a match. For example, the words ``THEIR" and "THERE" is a match, whereas the words ``DARK" and ``PRINCIPLE" are not phonetically matching. This section provides a base explanation of important phonetic algorithms relevant in this project.

\textbf{Soundex} - One of the earliest examples of a phonetic algorithm is known as Soundex. It was initially designed for phonetically indexing names alongside detection of transposed letters in spelling mistakes.  Soundex is one of the most famous example of a phonetic. This is due in part to its implementation into major database clients like MySQL\cite{mysql_soundex}, Oracle\cite{moved_2005} and PostgreSQL\cite{postgresql}.

\textbf{NYSIIS} - The New York State Identification and Intelligence System (NYSIIS) phonetic code was created for the phonetic matching of American names. The motivation for its inception was mostly due to the presence of Hispanic names in the American based databases (this was an aspect Soundex was known to have low accuracy with).

\textbf{Metaphone} - Metaphone was invented by Lawrence Philips in 1990\cite{philips1990hanging} in response to the deficiencies in Soundex. It improves on Soundex by including information around inconsistency and variation in English spelling in an attempt to create a more accurate phonetic representation.

\textbf{Levenshtein Distance} - Levenshtein distance is a string metric designed to measure the `distance' between two strings. It is merely the number of single-character edits (insertions, deletions or substitutions) required to reach the other string.

\textbf{Phonetic Vectors} - Phonetic Vectors is a unique addition to the chosen set. Created by Allison Parrish in 2017\cite{parrish2017poetic}, Phonetic vectors is as the name suggests the vectorisation of the phonetics of a word. Phonetic features are used in this work as a way to compare the similarity of phonemes. Phonemes are the phonetic elements that construct a word. For example, the word ``RING" translated into the phonemes \verb|/R IH NG/|.

% \subsection{Pretty Easy Privacy}
% \label{sec:pep}
% A very recent implementation of a word list can be found in Pretty Easy Privacy (\pep) implementation of TrustWords\cite{trustwords}. \pep is a data encryption system that utilises PGP encryption to provide E2EE on all common channels of communication such as email or SMS. The embedded design principles state that above all the systems should be easy to install, use and understand.

% \pep deals with the threat of MiTM attacks by having users compare the respective key fingerprints encoded as a set of words. Figure \ref{fig:trustwords} shows the \pep Android implementation of Trustwords. The users then authenticate the words on an OOB (Out of Band) channel such as a phone call or in-person communication. If both users decide the words match, they accept or decline respectively.

% \begin{figure}[h!]
%     \centering
%     \includegraphics[width=0.4\textwidth]{trustwords/trustword_handshake.jpg}
%     \caption{Trustword fingerprint verification}
%     \label{fig:trustwords}
% \end{figure}

% The unique aspect of TrustWords is its mapping of a single word to 16-bits. If compared to alternative literature, this is the highest number of bits-per-word seen. Full mappings (no duplication of words) would, therefore, require $2^{16}$ words in the dictionary. This number of words is arguably higher than most users' vocabulary. This deviation from the norm has not been currently backed up by research.

### 2019-08-31 Writeup/Paper/sections/design.tex
Changes; - Changes made to the precis paper structure

\subsection{Attack Design}
\label{sec:attackDesign}
The attack on Trustwords involves generating ``near-collision'' keys. Near-collision keys are keys composed of a set of words that are deemed a match by the similarity metric.
\\\\
As each combined key in Trustwords is an exclusive-or of both sides' public-key the attack is designed to target a single pair of users and requires recomputation for each attack target. Each pair is split into an "Uncontrolled" or "Controlled" key. Uncontrolled is the receiver of the communication, and, thus, the key cannot be altered. The Controlled key is the one we are attempting to impersonate. It is assumed that there is the ability to replace the Controlled key with the malicious option. The uncontrolled and controlled keys can be swapped around, thus, resulting in the possibility to intercept both directions of communication. This, however, requires performing the attack twice.
\\\\
When attacking, a similarity metric is used to compute a list of possibilities for each position in the target fingerprint. Completing these steps produces a list of fingerprints that can be inserted into a tool designed to hash a large number of keys and search for matches. This aspect of using an extensive list to search for keys massively reduces the complexity of the search.

In summary, the attack steps are:

\begin{enumerate}
    \item Compute all possible matches using a similarity metric on all words in a dictionary (Only needs performing once).

    \item Select a target and allocate "Uncontrolled" and "Controlled" key identification.

@@

The inspiration for the design of this tool was taken from a tool called Scallion\footnote{\url{https://github.com/lachesis/scallion}}. Scallion was designed by Richard Klafter and Eric Swanson and was used to demonstrate that 32-bit PGP key IDs were insufficient. To keep with the onion-based theme, the proposed tool is called `GreenOnion' and is a re-write of Scallion in C++. This language was chosen due to the well-understood efficiency benefits. The proposed tool differs from Scallion, most notably in its ability to concurrently search for a large number of keys, GreenOnion improves on this substantially. More implementation and experimental details are discussed in later chapters.

The tool should take two keys as parameters (Uncontrolled/Controlled) and a chosen similarity metric and produce a list of target key fingerprints. This list is then used as a search criterion when searching for keys. To utilise the parallel nature of the GPU to compute the hash of a large number of keys, the tool utilises a GPGPU (General-purpose computing on graphics processing units) framework. The chosen framework was OpenCL due to its support for the chosen language (C++) and platform (Linux). OpenCL allows the creation of code chunks referred to as "kernels" to be executed concurrently, this provides a massive speed increase compared to the sequential nature of the CPU.

>>>>

\subsection{Attack Design}
\label{sec:attackDesign}
The propossed attack on Trustwords involves generating ``near-collision'' keys. Near-collision keys are keys composed of a set of words that are deemed a match by the similarity metrics.
\\\\
As each combined key in Trustwords is an exclusive-or of both sides' public-key, therefore, the attack is designed to target a single pair of users and requires recomputation for each attack target pair. Each pair is split into an "Uncontrolled" or "Controlled" key. Uncontrolled is the receiver of the communication, and, thus, the key cannot be altered. The Controlled key is the one we are attempting to impersonate. It is assumed that there is the ability to replace the Controlled key with the malicious option. The uncontrolled and controlled keys can be swapped around, thus, resulting in the possibility to intercept both directions of communication. This, however, requires performing the attack twice.
\\\\
When attacking, a similarity metric is used to compute a list of possibilities for each position in the target fingerprint. Completing these steps produces a list of fingerprints that can be inserted into a tool designed to hash a large number of keys and search for matches. This aspect of using an extensive list to search for keys massively reduces the complexity of the search.

In summary, the attack steps are:
\begin{enumerate}
    \item Compute all possible matches using a similarity metric on all words in a dictionary (Only needs performing once).

    \item Select a target and allocate ``Uncontrolled" and ``Controlled" key identification.

@@

The inspiration for the design of this tool was taken from a tool called Scallion\footnote{\url{https://github.com/lachesis/scallion}}. The proposed tool is called `GreenOnion' and is a re-write of Scallion in C++.  The proposed tool differs from Scallion, most notably in its ability to concurrently search for a large number of keys. This is due to the unique addition of a bloom filter. A bloom filter is a probabilistic data structure that allows efficient checking for the presence of an element in a set. It is effectively a vast array of booleans that state whether an element is present. A pre-defined hashing algorithm decides the index in the array. This process is repeated several times with a set number of hashing algorithms ($k$) to populate the array of length ($m$). To check the presence of an element in the set means hashing the target value and checking the indices returned. This data structure, therefore, has a complexity of $O(k)$ regardless of the number of elements in the set. Due to the use of hashing algorithms, there is the possibility for collisions and, thus, the possibility of false-positives. The data structure, however, does not produce any false-negatives. If the levels of false-positives are controlled (by altering $k$ and $m$), the tool can search through a vast number of potential keys with minimal decrease in speed.

The tool should take two keys as parameters (Uncontrolled/Controlled) and a chosen similarity metric and produce a list of target key fingerprints. This list is then used as a search criterion when searching for keys. To utilise the parallel nature of the GPU to compute the hash of a large number of keys, the tool utilises a GPGPU (General-purpose computing on graphics processing units) framework. OpenCL allows the creation of code chunks referred to as "kernels" to be executed concurrently, this provides a massive speed increase compared to the sequential nature of the CPU.

### 2019-08-31 Writeup/Paper/sections/experimental_design.tex
Changes; - Changes made to the precis paper structure

\subsubsection*{Comparisons}
Each comparison contains a match-pair. These matches are generated by running the similarity metric over all the pairings in the Trustword dictionary; if the pair meets the criteria discussed in the previous section, it is determined a match. Potential matches are then sampled from their respective lists.

>>>>

\subsubsection*{Comparisons}
Each comparison contains a match-pair. These matches are generated by running the similarity metric over all the pairings in the Trustword dictionary. If two words are deemed a match they are added to their own list. Potential matches are then sampled from their respective lists.

### 2019-08-31 Writeup/Paper/sections/introduction.tex
Changes; - Added inital paper content
Changes; - Changes made to the precis paper structure

The increasing use of public-key cryptography by instant messaging and secure email means ensuring confidentiality is an ever more important task.
One of the most significant risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack. A MiTM attack involves an attacker impersonating one or both sides of a connection. MiTM attacks can entirely circumvent the encryption as it allows an attacker to read all of the encrypted data.
A countermeasure for the threat of MiTM attacks is the verification of each parties’ fingerprint. A fingerprint is a small string of characters that is unique to each key and, thus, can be used to identify.
Fingerprints can come in several different encodings such as Hexadecimal, words, and even procedurally generated avatars. Previous research has shown that the average human can only hold around 7-digits ($\pm 2$) worth of data in their working memory\cite{miller1956magical}. Consequently, this makes the designing of user-friendly schemes a task of utmost importance.
Humans are commonly considered the most vulnerable part of any computer system. Solutions, therefore, have been proposed to remove this manual verification with examples such as PGP's web-of-trust\cite{callas1998openpgp} and Namecoin\cite{kalodner2015empirical}. However, these suffer from user adoption due to perceived complexity. Manual verification is, consequently, left in a difficult position, due in part to proposed solutions needing to sacrifice either usability or security. Therefore, research into improving or creating a more secure and user-friendly fingerprint encoding remains an important task.
One proposed user-friendly scheme is Pretty Easy Privacy's (p$\equiv$p) "Trustwords". Trustwords is an implementation of word fingerprint mapping with a emphasis on usability. This increase in usability is achieved by the user comparing a reduced number of words. This usability boost, however, comes at a cost of a much larger word list.
This report evaluates the security and feasibility of attacking Trustwords. The motivation for this is due in part to the shortage of justification behind the size and features of the chosen wordlist.

>>>>

The increasing use of public-key cryptography by instant messaging and secure email means ensuring confidentiality is an ever more important task.

One of the most significant risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack. A MiTM attack involves an attacker impersonating one or both sides of a connection. MiTM attacks can entirely circumvent the encryption as it allows an attacker to read all of the encrypted data. A countermeasure for the threat of MiTM attacks is the verification of each parties’ fingerprint. A fingerprint is a small string of characters that is unique to each key and, thus, can be used to identify.

Fingerprints can come in several different encodings such as Hexadecimal, words, and even procedurally generated avatars. Previous research has shown that the average human can only hold around 7-digits ($\pm 2$) worth of data in their working memory\cite{miller1956magical}. Consequently, this makes the designing of user-friendly schemes a task of utmost importance.

Humans are commonly considered the most vulnerable part of any computer system. Solutions, therefore, have been proposed to remove this manual verification with examples such as PGP's web-of-trust\cite{callas1998openpgp} and Namecoin\cite{kalodner2015empirical}. However, these suffer from user adoption due to perceived complexity. Manual verification is, consequently, left in a difficult position, due in part to proposed solutions needing to sacrifice either usability or security. Therefore, research into improving or creating a more secure and user-friendly fingerprint encoding remains an important task.

One proposed user-friendly scheme is Pretty Easy Privacy's (p$\equiv$p) ``Trustwords''. Trustwords is an implementation of word fingerprint mapping with a emphasis on usability. This increase in usability is achieved by the user comparing a reduced number of words. This usability boost, however, comes at a cost of a much larger word list ($2^{16}$ words).

The project aims to assess the security of p$\equiv$p’s minimum recommendation of four Trustwords\cite{trustwordsHandshake}. Moreover, issues with the dictionary's design, such as the presence of homophones and dual-mapping of words show the potential for possible vulnerabilities.
